---
output: word_document
geometry: margin=1in
fontsize: 11pt
---
```{r Initialization, include=FALSE}
#Initialization code - includes Libraries, Functions, Data Load, Globals
library(sp)
library(rgdal)
library(RColorBrewer)
library(gridExtra)
library(lattice)
library(maps)
library(maptools)
library(rgeos)
library(png)
library(grid)
library(Cairo)
library(data.table)
library(grid)
library(gridExtra)
library(splancs)

#------------------------------
#Globals & Settings
#------------------------------
wd = "C:\\Scratch"
setwd(wd) 

#Lat/Lon extents for maps (for easy reference)
#(w,e)
w_e = c(80,88.25)

#(s,n)
s_n=c(26,30.5)

buffer = 50

iterations = 1000

#------------------------------
#Define functions
#------------------------------
commaUS <- function(x) sprintf("$%s", formatC(x, format="fg",  
big.mark = ","))



theme.novpadding <-
    list(layout.heights =
         list(top.padding = -10,
              main.key.padding = 0,
              key.axis.padding = 0,
              axis.xlab.padding = 0,
              xlab.key.padding = 0,
              key.sub.padding = 0,
              bottom.padding = -10),
         layout.widths =
         list(left.padding = 0,
              key.ylab.padding = 0,
              ylab.axis.padding = 0,
              axis.key.padding = 0,
              right.padding = 0))
#------------------------------
#Load in Data
#------------------------------
GIS_csv = file.path("Nepal_AidData","NPL_geocoded_projectLocations.csv")
path = normalizePath(GIS_csv)
GIS_raw_data = read.csv(path,stringsAsFactors = FALSE)

ADM_GIS = file.path("Nepal_AidData\\GIS","NP_ADM4.shp")
ADM_path = normalizePath(ADM_GIS)
ADM_shp = readShapePoly(ADM_path)

AMP_data_csv = file.path("Nepal_AidData","NPL_AMP_projects.csv")
AMP_path = normalizePath(AMP_data_csv)
AMP_raw_data = read.csv(AMP_path,stringsAsFactors = FALSE)

GIS_AMP_database = merge(GIS_raw_data, AMP_raw_data, by="project_ID")

Geo_db <- GIS_AMP_database
Geo_db <- Geo_db[rowSums(is.na(Geo_db[,c("latitude","longitude")]))==0,]

Geo_db = SpatialPointsDataFrame(list(Geo_db$longitude,Geo_db$latitude), Geo_db)

ADM_0 = gUnaryUnion(ADM_shp)

agex <- AMP_raw_data[grep("AGRICULTURE", AMP_raw_data$amp_sector_name),]
geo_agex <- Geo_db[grep("AGRICULTURE", Geo_db@data$amp_sector_name),]

```



STATUS: Draft - 0.1
GOAL: Academic Publication
TARGET JOURNAL: Climate Risk Management (via Agriculture angle)
AIDDATA BENEFIT: Method for quantifying aid at relevant scales of analysis; quantify uncertainty associated with our precision codes; example of Decision Support capabilities using our data; Value of Information (what our information is worth in terms of reducing uncertainty; what collecting more would net).
CO-AUTHOR OPPORTUNITY: GIS Methods piece; outreach to Data team.  Opportunities include writing portions of literature review, writing code to operate in other countries, effeciency gains in code for operational use, extending to new topical areas.


The Value of Spatial Information for Decision Support Systems : International Agricultural Aid in Nepal
========================================================
Daniel Miller Runfola, Albert Decatur, Doug Nicholson, Scott Stewart

Abstract
-----------------------------
Spatial decision support systems (SDSS) provide decision makers with the capability to transparently identify statistically optimal strategies for the allocation of limited resources.  Many organizations have begun mapping current resource allocations in order to take advantage of these systems.  Here, we build on previously developed, Monte-carlo approaches to quantifying spatial uncertainty to illustrate the value of this information in the context of multi-scalar precisional uncertainty.  A case study using a simplified agricultural SDSS and spatially-explicit information on interational aid for agriculture in Nepal is presented.  Findings suggest that...

Key Words
-----------------------------
Uncertainty, Decision Support, International Aid, Development Fianance


1.0 Introduction
-----------------------------
**1.1 Research Objective**
**1.2 Definitions**
**1.3 Literature Review**
[Forthcoming]


2.0 Data & Methods
-----------------------------
**2.1 Data**
Two data sources for Nepal are used in this analysis: georeferenced information on international aid projects for agricultural improvement {AidData; Tierney et al. 2011} and aggregate yield ratio information (a ratio of actual production compared to potential agro-climatic production) for rice {IIASA 2012}. The AidData information contains records for `r nrow(agex)` agricultural projects that operate in Nepal, of which `r nrow(agex[agex$geocoded == 1,])` have location information in `r nrow(geo_agex@data)` identified locations. 


```{r figure1, echo=FALSE}
par(mar=c(0,0,0,0),mfrow=c(1,1), pty="s")
#spplot(geo_agex, c("precision_code"), draw=T, cuts=1, axes=TRUE, scales=list(draw=TRUE), auto.key=FALSE,legendEntries=c("Agricultural project location"),  xlab=expression("Longitude"^o), ylab=expression("Latitude"^o), col.regions=c("Black"), cex=0.3, sp.layout=list("sp.polygons",ADM_0), par.settings=theme.novpadding)

spplot(geo_agex,c("precision_code"),cuts=1,col.regions=c("black"),legendEntries=c("Agricultural Aid Project"),cex=0.5, xlab=expression("Longitude"^o), ylab=expression("Latitude"^o), sp.layout=list("sp.polygons",ADM_0),axes=TRUE, scales=list(draw=TRUE), xlim=w_e, ylim=s_n)

```
```
Figure 1. Location of Agricultural Projects in Nepal.
```





**2.2 Methods**
**2.2.1 Methods Overview**

To assess the value of spatial information, three models are produced:
(A) A model which optimizes based on no spatial information (uniform allocation across all administrative units);
(B) A model which optimizes based on only spatial information on yield ratio information (no information on where existing agricultural aid projects are); and
(C) A model with spatial - but uncertain - information on the location of current agricultural aid.

The result of each optimization is compared to the simplified "ideal" map of aid allocation, in which aid is allocated proportionally to where agricultural gains are biophyiscally possible.^[While an illustrative example, this approach could easily be complicated in practice by further including information on - for example - existing infrastructure (i.e., roads), population, crime, or any other number of spatial data layers that may be relevant for agricultural project allocation.]  This comparison is done both spatially - i.e., how different optimization strategies differ from optimal across different administrative units - as well as in terms of total quantity.  A few simple assumptions are made to facilitate this comparison.  First, we assume that Nepal is doubling the total amount of agricultural aid in the country - from 552 million to 1.14 billion dollars.  This gives past and future decisions equal weight.  Second, we assume that existing projects cannot be re-allocated - that is, decision makers cannot remove funding from an area, only allocate new funding.  Third, we assume that one dollar of agricultural aid would have equal impact in all administrative units, and that this impact is linear in nature (that is, an additional dollar will always yield the same increase in agricultural productivity irrespective of the current agricultural productivity). 

**2.2.2 Estimating Existing Aid**
In order to estimate the current amount of aid present for agriculture in each administrative unit, the generic monte carlo procedure suggested in {Jeroen et al. 2003} is modified to account for the multi-scalar precisional uncertainties in the AidData database.  For example, many of the aid allocation projects present in the database are known to be allocated at a single project point (no precisional uncertainty), while other projects are only known at a given administrative unit (i.e., "an unknown location within a district") or simply known to be present somewhere witin the country.  Another type of error is for those projects which are split across multiple points.  For example, agricultural projects to drill wells for irrigation may have a known total funding amount and perfect spatial information, but the geographic allocation of this funding across points is uncertain.

To quantify the impact of these interacting uncertainties, multiple equiprobable surfaces of aid allocation are created following the following, iterative procedure:
(1) For projects which have precisional uncertainty (i.e., not georefereced to a single point), a random number of points are created at random locations across the unit of analysis the project is known to be located in.[^Rarely, aid projects are aggregated to administrative units for which no geographic representation currently exists - i.e., a national park.  In these cases, an arbitrary buffer of `r buffer` kilometers is used to allocate aid.]
(2) The total aid dollars for a project are randomly allocated across all geographic units they exist at.
(3) Aid dollars are aggregated to an approximately 9km resolution (.0833 degree) raster grid (chosen to match the resolution of available agriculture information).

This procedure is repeated `r iterations` times, and summary statistics for all cells are calculated across all iterations (producing five surfaces reporting mean agricultural aid dollars, standard deviation, range, min, and max).  A single example iteration is outlined in figure 2(a) through figure 2(f).  Figure 2a shows a single spatial allocation of possible aid project sites, representing step (1) above.

```{r figure2, echo=FALSE}
par(mar=c(0,0,0,0),mfrow=c(1,1), pty="s")
#spplot(geo_agex, c("precision_code"), draw=T, cuts=1, axes=TRUE, scales=list(draw=TRUE), auto.key=FALSE,legendEntries=c("Agricultural project location"),  xlab=expression("Longitude"^o), ylab=expression("Latitude"^o), col.regions=c("Black"), cex=0.3, sp.layout=list("sp.polygons",ADM_0), par.settings=theme.novpadding)

spplot(geo_agex,c("precision_code"),cuts=1,col.regions=c("black"),legendEntries=c("Agricultural Aid Project"),cex=0.5, xlab=expression("Longitude"^o), ylab=expression("Latitude"^o), sp.layout=list("sp.polygons",ADM_0),axes=TRUE, scales=list(draw=TRUE), xlim=w_e, ylim=s_n)

```
```
Figure 2. Example of a single Monte Carlo iteration: allocation of project locations.
```


**2.2.3 Modeling Strategies for Allocating New Aid**
Three modeling strategies are followed, each with an increasing amount of spatial information.  


In model *A*, a theoretical decision maker allocates new aid for each administrative unit according to the following equation:

$\frac{1}{n} \sum_{i=i}^{n} x_{i}$




3.0 Results
-----------------------------

4.0 Discussion
-----------------------------

5.0 Conclusion
-----------------------------

Acknowledgements
-----------------------------

Literature
-----------------------------

Tables
-----------------------------

Figures
-----------------------------








